---
title: "Homework 5"
author: "Yishi Wang"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggplot2)

knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

## Problem 1
Build a function that draw n birthdays and return true if there are duplicates. We use set.seed() for reproducibility:
```{r}
set.seed(100)

has_duplicate_bday = function(n) {
  bdays = sample(1:365, n, replace = TRUE)
  return(any(duplicated(bdays)))
}
```

Run simulations and save results as data frame:
```{r}
group_sizes = 2:50
bd_simulation = sapply(group_sizes, function(n) mean(replicate(10000, has_duplicate_bday(n))))
bd_simulation_df = data.frame(n = group_sizes, prob = bd_simulation)
```

Plot our simulation result:
```{r dpi=300}
bd_simulation_df |>
  ggplot(aes(x = n, y = prob)) +
  geom_line() +
  geom_point() +
  labs(
    x = "group size",
    y = "P(at least one shared birthday)",
    title = "Probability that duplicated birthday exist in n people"
  ) +
  theme_minimal()
```

From the plot above, we see the probability that duplicated birthday exist in a group of people increases rapidly (almost linearly) as group size increases. It is very likely that a group of 50 people have shared birth dates.

## Problem 2
Build a function that simulates power in one-sample t-test:
```{r}
n = 30
sigma = 5
mus = 0:6

explore_test_power = function(mu) {
  replicate(5000, {
    x = rnorm(n, mean = mu, sd = sigma)
    tt = t.test(x, mu = 0)
    broom::tidy(tt)[c("estimate", "p.value")]
  }, simplify = FALSE) |>
    bind_rows(.id = "rep") |>
    mutate(rep = as.integer(rep))
}
```

Run simulations and save results as a data frame:
```{r}
set.seed(100)
sim_results <- lapply(mus, explore_test_power) |>
  setNames(as.character(mus))

summary_df <- bind_rows(
  lapply(names(sim_results), function(mu) {
    dat <- sim_results[[mu]]
    dat$mu_true <- as.numeric(mu)
    dat
  })) |>
  group_by(mu_true) |>
  summarise(
    power = mean(p.value < 0.05),
    mean_est_all = mean(estimate),
    mean_est_rejected = if (sum(p.value < 0.05) > 0) mean(estimate[p.value < 0.05]) else NA_real_,
    n_rejected = sum(p.value < 0.05),
    .groups = "drop"
  )
```

Make a plot showing the proportion of times the null was rejected:
```{r dpi=300}
ggplot(summary_df, aes(x = mu_true, y = power)) +
  geom_line() + geom_point() +
  labs(x = "true mu", 
       y = "power",
       title = "Power vs True Mu")
```

From the plot above, we see the power increases as the true value of $\mu$ increases, because large effect sizes are easily detected by t-test. Now we
make a plot to show the average estimate of all and rejected $\hat\mu$:
```{r dpi=300}
ggplot(summary_df, aes(x = mu_true)) +
  geom_line(aes(y = mean_est_all), linetype = "solid") +
  geom_point(aes(y = mean_est_all)) +
  geom_line(aes(y = mean_est_rejected), linetype = "dashed") +
  geom_point(aes(y = mean_est_rejected)) +
  labs(x = "true mu", y = "avg sample mean",
       caption = "solid = all  dashed = rejected",
       title = "Mean Sample Average vs True Mu")
```

From the plot above, we see the difference between $\hat\mu$ for which the null is rejected differs and true $\mu$ gets smaller as true $\mu$ gets bigger. Selecting tests with rejected null keeps only the large sample means, but when almost every sample gets rejected (power approaches 1), the selection bias disappears.

## Problem 3
Import and check the data:
```{r}
homicides_df = read_csv("local_files/homicide-data.csv") |>
  janitor::clean_names()
```
This dataset is gathered by Washington Post and has data on homicides in 50 large U.S. cities, main variable names include:

**uid** — unique record id.

**reported_date** — date the homicide was reported (YYYYMMDD).

**victim_last, victim_first** — victim name parts.

**victim_race** — victim race/ethnicity.

**victim_age** — victim age.

**victim_sex** — sex/gender.

**city, state** — city and 2-letter state.

**lat, lon** — latitude/longitude.

**disposition** — case disposition.

Next, we create city_state variable and summarize cities to get total number of homicides and the number of unsolved homicides:
```{r}
homicides_df = homicides_df |>
  mutate(city_state = paste0(city, ", ", state))

city_counts = homicides_df |>
  group_by(city_state) |>
  summarise(
    total_homicides = n(),
    unsolved_homicides = sum(disposition %in% c("Closed without arrest", "Open/No arrest"), na.rm = TRUE),
    .groups = "drop"
  )
```

First, filter out the city of Baltimore and use prop.test to estimate the proportion of unsolved homicides:
```{r}
baltimore_counts = city_counts |>
  filter(city_state == "Baltimore, MD")

baltimore_prop_test = prop.test(
  x = baltimore_counts$unsolved_homicides[1],
  n = baltimore_counts$total_homicides[1]
)
```

Save the result from our test above:
```{r}
saveRDS(baltimore_prop_test, file = "local_files/baltimore_props.rds", compress = "xz")
```

Tidy and pull the estimated proportion and confidence intervals:
```{r}
baltimore_tidy = broom::tidy(baltimore_prop_test)

# The estimated proportion is:
baltimore_tidy$estimate

# The lower CI is:
baltimore_tidy$conf.low

# The upper CI is:
baltimore_tidy$conf.high
```

Now we run prop.test for all cities similar to what we did on Baltimore, and return a tidy data frame with purrr package and unnest() 
```{r}
city_props = city_counts |>
  filter(total_homicides > 0) |>
  mutate(
    prop_test_obj = purrr::map2(unsolved_homicides, total_homicides, ~ prop.test(.x, .y)),
    tidy_test = purrr::map(prop_test_obj, broom::tidy)
  ) |>
  unnest(tidy_test) |>
  select(
    city_state,
    estimate,
    conf.low,
    conf.high,
  )

city_props |> 
  knitr::kable()
```

Lastly, plot our results above:
```{r fig.width=8, fig.height=8, dpi=300}
city_props_plot = city_props |>
  mutate(city_state = fct_reorder(city_state, estimate))

unsolved_prop_plot = ggplot(city_props_plot, aes(x = city_state, y = estimate)) +
  geom_point(size = 2) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +
  coord_flip() +
  labs(
    x = "city",
    y = "estimated proportion unsolved",
    title = "Proportion of Unsolved Homicides by City",
    caption = "Points = estimate   bars = 95% CI"
  ) +
  theme_minimal()

print(unsolved_prop_plot)
```


